Transforming Trash as an Artistic Act

### 2016-01-17 chapter4_final.tex
Print copy preperations, proof reading
Former-commit-id: c4e7e2d53515fdd32b70a48f44d38c9e790b69b9

In this chapter thesis project and its development progress will be explained in detail in the light of what has been discussed in previous chapters. The project will be examined in various dimensions. The relationship will be established between the previously mentioned artworks. How this project is related to the discussed concepts and how it can be positioned among the previous works will be explained.

Firstly, brief statement and description of the work is given. Secondly, the development of the project and experiments are explained. Lastly, the final body of the work with its parts is presented.

%
%
% TODO: Be careful on tenses.
\section{The Statement}
Refuse is part of people’s consumption practices. It is a very common concept from developed cities to rural areas, from modern societies to ancient societies. It is not wanted, thrown away, kept away. It is considered as useless, disgusting, abject, unattractive, and harmful to the people and nature. Contrary to the common approach as \cite{thompson1979rubbish}, author of the book Rubbish Theory, stated \quotes{one man’s trash is another man’s treasure}.

@@

Through this project trash were re-imagined (re-considered, re-thought), new possibilities and alternatives were seek. As it is discarded, missed potentials were discovered.  This project aims to find new potentials on trash.

The purpose is not to build an object that is produced from discarded material to watch it from distance. The important thing is to interact with it. It is already discarded and general behavior avoid from it. This work must change it. It must call the viewer to interact with it. Trash must be accepted by them. This project asks people to readmit the trash to the their life. For this reason notebooks are given away free at various places where people visit frequently. By doing that once thrown materials would find a place with new purposes in the community.

%
%
\section{Development Progress of the Project}
In order to have an insight of this project and realize artistic intentions, it is important to look at the process that reveals the path to the final work and critical decisions that were necessary to the completion of the project.

@@

% This progress also reveals that how this work can be positioned among the other works.

@@

% [SODA CAPS]
I remember from my childhood that we were collecting soda caps and were playing with them. We were looking everywhere to collect them. Some caps were found less and they worth more in the game. We put them on the railways to make them flat. After the train had passed we had perfect flat caps for our games. At that time, it was not waste for us. It has a value and part of our games and indispensable source of fun. The notion of playing with it can be related with the Picasso’s collage work \quotes{Still Life with Chair Caning} mentioned previously. He pasted letters of the French word \quotes{JOU} means \quotes{game} signifying that playing with the materials. We were already playing with trash while we were child. Therefore, it can be better understand that why Picasso said that \quotes{every child is an artist}.

% [VASVIYE HOCA]
Another important memory is from my primary school teacher. When I was at third class, our teacher wanted us to bring colorful papers to make something (whatever it is I do not remember now). The day after we brought some colorful paper that was bought from stationery store. Our papers’ shape and color looked same. However her papers were different in every dimension. They were cut out from packages and advertisements. I remember clearly that she suggested that same for us. \quotes{Do not throw out packages, look for the useful parts and keep them to use later.} Whenever I am going to throw something away, this comes to my mind. Can it be reused? Every time it is not obvious that how can it be useful, but at some point it becomes an indispensable practice. Further, this practice is also support to the bricolage methodology. Keeping many things makes easier to build (or produce) whatever at hand.

@@

I am very uncomfortable with the act of discarding. I often think that by discarding objects, I am missing their potentials. I must find another use which is more effective than throwing away. At least I can give them to someone else who can use it for own purposes. Seeing papers on the waste bin is a disappointment for me. As artist Aaron Kramer argues that \quotes{[t]rash is the failure of imagination} \citep[as cited in][]{meyer2007turning}. There is a lot of effort to produce objects and it seems that all this efforts are wasted. What I mention is not related about ultimate productivity. It is more close to being thoughtful, and taking responsibility of tools, items and objects that we are using. Rather than throwing out, creating a way that all are have chance to live together is much more close to my perception. On the contrary, sometimes I became overwhelmed because of the significant amount of leftover from daily activities and in order to move on I need to leave them behind. It is the result of inevitable rhythm of modern life style. Consuming and wasting is part of people’s daily routine and most of the time they are not aware of what they consume, and what they discard. In the scope of this thesis subject, saving objects rather them getting rid of them is an effective way of breaking the routines. The use of that object can be decided in the future.

% [encouraging point]
One of the encouraging factor of working on trash is to explore something different that others cannot see. In other words, it contains eureka moments. Because it is ignored by discarding and therefore undiscovered. It waits us to be discovered and possibly (as Picasso’s explanation on his work ’Bull’s Head’\footnote{See epigraph in the beginning of Chapter 3}) discovered again and again.

% [EXPERIMENTS AT HOME]
My remarkable interaction with trash is from recent years. Long years I stayed at dormitory during when I was high school and university while undergraduate. Dormitory is place that living spaces is shared with many people. Therefore there is a limited place for you and your possessions. There is no space for unused items. After I graduated from university I started to living a rented house. This give me to opportunity to save the materials to use them later. My very first intention is to use them on my painting. Instead of white blank canvas I want to experiment with different surfaces. I collected boxes and cart boards from my friends. My aim is to turn discard materials to the different things. Here is some of the my previous works.

@@

After turning discarded materials to different uses I began to collect them more eagerly. At the same time I also gave what I did to my friends as a birthday gift. I was experimenting with the material and at some point with the help of my mother I made my first notebook covered by Starbucks package. Papers inside of it are blank mimeograph paper which different than white paper. Particularly purpose is to move away from white paper because I feel uncomfortable when using it. It looks to me very bright. Therefore mimeograph paper is better for me. At that time it was a great victory for me. I started to carry it and use it every possible place. I show them to people and some of them supported to me giving their discard. One of them is my supervisor. He started to collect packages for me. After turning his packages to notebooks, he suggested me to go deeper in this subject and make my thesis project from them. Thereafter what I need is more discarded material.

@@

% NEW PART [collecting] 
\subsection{Collecting}
Collecting discarded objects has carried out through the efforts of mine and people around me. Lars Eighner’s instructions helped me a lot when collecting objects. He writes that:

\begin{quote}
Eating safely from the Dumpsters involves three principles, using the senses and common sense to evaluate the condition of the found materials, knowing the Dumpsters of a given area and checking them regularly, and seeking always to answer the question \singlequotes{Why was this discarded?} \citep[as cited in][6]{strasser1999waste}.
\end{quote}

@@

Moreover there are other rules that define the frame of collecting act. Paper or package must be trash or will be trashed. For this project anything must not intentionally bought to use as trash. In other words first rule is to consume less and generate less trash. Lots of trash is already produced every time everywhere. Aim is to reach them.

% OWNER
\quotes{The question of who owns these discards is not trivial} \citep{zimring2012encyclopedia}. This question makes me feel uncomfortable because of thinking that stealing that someone elses’ objects. Particularly I felt it when collecting leaved out papers from Bilkent Computer Laboratory. You already know that it owned by somebody else and after a while owner can take them. I solved this issue checking the leaved papers in every couples hours. If it stays there I take them. Further I encountered that If nobody takes them for a long time servants collect them.

@@

% CONTRIBUTIONS OF PEOPLE
As mentioned previously many people contributed to the project by collecting. Whenever I meet my friends I ask them to save their discard for the project. The importance of collaboration with them can be understood the term \quotes{citizen science} which is \quotes{defined as scientific research performed in part or in whole by volunteers who are not professional scientists} \citep{robson2012using}. I need to note that this is not a scientific project but the notion belongs there. The value of it lies in the collective intelligence of the contributors. In the scope of this thesis the term can be interpreted as people bring their own creativity and interpretation to the project by collecting the items that they reach and select. When the number of contributors are increased the diversity of the collected material will be increased and this will echoed in the notebooks.

It can be viewed as a co-authored work. Because many people helped me to collect the papers and also many people leave their trail to the paper and also they bring meaning to the them. After production of papers I also want them to continue as co-authored process. In other words many people will be part of this project. As people see different things on objects and bring them to me. the variety of things is increased in time. Because people sad me that \quotes{I thought that you can use it. Do you collect also these?}. They bring their unique approach.

@@

% What type of activities generates waste.
Beyond social status it reveals activities that generate such a waste.It can be also extracted that most of the trashed papers are thrown away from feeding activity. On the other hand vastness of some types of disposables reveal the dark side of the enterprises such as Starbucks and Burger King. Few people thought that after seeing the project I only collected Starbucks packets. There is no such a limitation but as a result of my habitat the number of them suppress the others.

@@

% TODO see figure appendix.
Collected items are stored in my room for a long time. Most of them are not belong to me. Here the detailed information of the collected materials including the place where it is collected and their previous usage are listed.

\textbf{Burger King packages} (Figure \ref{fig:BurgerKingBag}) are collected from the office where I work. At launch time, some of the colleagues order their meal from Burger King which is a global chain of fast food restaurant. Meals are delivered to the office with paper packages. The package of the meal is collected and used as a cover of notebooks.

Inside the bag one paper package is used for French fries which is oily and paper is very good to absorb oil. Some people think that it is disgusting to use it for other purposes because of oil onto it. But immediately I replied them \quotes{you just eat that oily fires so what is wrong with it?} At the same time what \citet[5]{strasser1999waste} said comes to my mind; \quotes{Nothing is inherently trash}. She refers to resolution on dirt of \citet[36]{douglas2003purity}:

@@

\textbf{Starbucks packages} are collected from my supervisor, office where I work and my friend. Particularly my friend saved and collected them from her friends. Few of them collected from waste bins. They are used as a cover of notebooks. There are different types of packets from this category. As claimed by the company all of them are build from recycled ---precisely downcycled--- paper. There is also one category of object that is not paper. It is package of coffee beans. It is airtight and waterproof. It is very solid and strong. It has well graphic designs with different color and illustration.

\textbf{Modshifters papers} are used to cover the dinner table. Beginning of the table there is a big roll of paper to cover the table. At the end of the day used parts are cut out and thrown away. One of my visit I stayed there at a late time and catch the waiter when collecting the used papers and preparing tables to the next customer. I asked kindly him to give me these papers and he accepted. They are used as a cover of notebooks. All of them contains track of people who sit down there. I do not know any of them and I am not sure that the paper that they are laid down on their dinner turned to such a thing.

@@

\textbf{Order slips} are collected from bank office located at Technopolis of Middle East Technical University. They are used to record the order of customers. I realized that back of them are empty and appropriate for writing. Further there many others are waiting on the bin. After this experiment I really convinced that I am really looking for trash everywhere every time. The act of collecting spread through the every moment of life.

@@

% NEW PART [transforming]
\subsection{Transformation}
At one side while paper is being collected, on the other side I started to make experiments with the material. I questioned what else can be made with these collected materials and tried to cover (wrap out) objects and plants in the public. However these do not give the result what I want. People cannot interact with these and possibly they will turn to trash again. For that reason the first approach which is making notebooks are more appropriate to accomplish the purpose.

@@

% [notebooks]
There are two main parts of notebooks; cover and pages. Papers that are mostly blank(empty) are used inside of the notebooks for writing. Others that are solid and colorful are used as a cover for notebooks. Mixing (juxtaposing) materials is main method in the production process. I combine different papers and packages. I try to generate every possible combination. The purpose is to create as many alternatives as I can because trash is ignored and people do not give it new alternatives. Therefore these notebooks provide many alternatives to show that otherwise is possible.

@@

% TODO starbucks defterlerinin hep yapıştırılmış olmasını da eklemek gerekli.

% TODO muhakkak tekrar kontrol etmek gerekli.
Size of the paper used inside of the notebooks are mostly equals to the size of an A4 sheet. Usage of papers differs among the notebooks. Especially for the papers taken from Varuna Gezgin written part is folded inward by hiding it. However for the ones collected from Bilkent Computer Laboratory written side of it leaved open. Their content comes from various students whose departments are also various. In other words offers great diversity and they are mixed to allow different interpretations. Further it provides what students from other departments works on. What other students work on can be understand from these.

There are different dimensions of notebooks. Most of them are half of A4 (210mm x 297mm) or quarter of it. Collected papers dimensions shape the size of the notebooks.

% TODO Dimensions of notebooks. most of them half of A4 or quarte of it. böyle bir fotoğraf eklenmeli. farklı boyutlarda farklı şekillerde ve renklerde.

@@

While transforming collected materials my method can be explained with the term upcycling mentioned at the end of the Chapter 1. Further I tried to use objects already at hand. In other words I followed the bricolage technique. However I have to do buy some tools for my work. Therefore it is hard to consider it as bricolage but my intention is to follow it.

Moreover I have reviewed other notebook projects. Notable ones are The Sketchbook Project and myDetour from Moleskine. In these projects anyone who bought the notebooks from them can participate and send their notebooks filled with illustrations and paintings. These projects can be seen as crowd-sourced library. Further the pages of the notebooks are available at online. After checking sketches in these notebooks, the diversity of them surprised me a lot. It forced me think that the transformation process is not limited with the production of notebooks from trash. After notebooks are given to the people, their transformation will continue. Therefore I think the process of transformation is a continuous act.

@@

% NEW PART
\subsection{Demonstration}
After producing notebooks the most important point is to present and reach to people. Main question is how to give them to people. Starting point is to stack the notebooks onto the each other, and there are other ideas like putting them inside a waste bin to re-do the action of throwing away. After ward somehow they should understand that these notebooks can be taken. I set up models and tried some installations of them.

@@

They were early interpretations and ideas. I imagined an exhibition environment to place notebooks. Different placements were tried. All of them looked like a sculpture. I have inserted no special meaning as sculpture to this objects. I also afraid that people will not interact with them and look them away. My purpose is give people. Aim is to spread the idea by making something useful from trash. Increase diversity, encourage people to embrace the trash.

@@

% TODO boxes and wooden cases
To hold notebooks together I made several small boxes from discarded card boxes. Notebooks were put inside of them.

In the production phase of the boxes, I discovered that my folded notebooks are fit into the many card boxes. In other words dimensions of them are not very different. This property of items make my production more easier. It shows that industrial system supports the reuse of the object with less effort in this way.

@@

%
%
\section{Parts of the Final Work}
As also can be understood by the previous section the project introduced in this thesis have different parts. Each part support the other. They are connected with each other. Different parts lead different inquiries.

%
%
\subsection{Notebooks}

Notebooks have been used for different purposes such as drawing, writing and recording, ideas or memorandum by many artists, scientists, and thinkers through the ages. Further same exist for me and people around me. Although the digital alternatives of it paper maintains its place in the community.

@@

% Meaning of stories behind the process
There are different stories behind the objects. I try to record their stories (by photography and taking notes) but some of them were not possible. However I still use them in the work because even if I missed their story, with their materiality reveals their history. It still has a history but needs to rewritten. Maybe forgetting all the history maybe creates different notebook.

@@

Giving name is another hard topic for me. However point out the story behind the notebooks and properties of it. These names gives clues about how they are combined what they represent, defines the aspects of them. Name of the paper can be related with this place and related with the meaning. These names captures the memory of notebooks. However for one series I borrowed the idea of nonsense word generation from the Kurt Schwitters. Similar to the the production of the nonsense word merz I decided to name some of the notebooks a nonsense word. There are notebook series that have a name. Here are the name of notebooks and their story:

@@

\textbf{Daro.} For some of the notebooks it is not obvious to name them. At this point approach of Kurt Schwitters provides me a way. He generated his frequently used word \quotes{Merz} from an advertisement. It is a nonsense word and used in most of his works. I followed what he did and generated a word from the papers that used inside of the notebooks. I use the word \quotes{Daro} extracted from \quotes{Vardaroma Dondurmacısı} for the ones whose story are not obvious and significant.

%
%
\subsection{Exhibition}
Especially it is the most hard topic in the development of the thesis. To find an appropriate place for the notebooks is great challenge for me.

\textbf{Public space.} The reason why select library is there are small papers to write down the location of books. Libraries are places that things are reused. Books are used by many people. It is place of sharing culture. Books are shared by all the other peoples. Public space. It is place that students from different department use the same place.

Notebooks are placed where people frequently visit such as computer desks.

\begin{figure}[!tbp]
  \begin{minipage}[b]{0.48\textwidth}
    \includegraphics[width=\textwidth]{project_graphics/bilkent1.jpg}
    \caption{Library}
    \label{fig:BilkentLibrary}
  \end{minipage}
  \hfill
  \begin{minipage}[b]{0.48\textwidth}
    \includegraphics[width=\textwidth]{project_graphics/bilkent2.jpg}
    \caption{Computer Laboratory}
    \label{fig:BilkentLaboratory}
  \end{minipage}
\end{figure}

% su anda sadece bilkente yerlestirildi. Odtuye daha koyamadim. Neden koymadim. Nicin koymadim. Koymali miyim?

@@

% Selected places are frequently visited by people. Shared places.

% Torun
Through this project another exhibition alternative is thought. In the notable museums and art galleries there are gift shops that people can buy souvenirs. Often they are mass produced imitations and replicas of well-known artists work embedded various objects such as cover on the notebooks. At this point I image an art space you can buy these types of items freely. In other words alternative to the existing approach this types of gadgets. Another alternative interpretation can be giving away different that the industrial items such as notebooks from trash as in my project. For this type of approach Torun is a great place to show my notebooks and give them away. Torun describe itself as a place for free place for sharing art. Open to everybody. Makes open call to the any artist. There are no security cameras on this gallery therefore no one is watching you at there. This idea shared with Torun and they accepted it. However Torun will opened after 14th of the January currently I just took some photos at there.

I used wooden market boxes to lift up them from ground to more reachable height. It is hole body of stand to put notebooks. Notebooks are placed on the top box and below them I just put the collected items to provide for those who want to make their own notebook. People can get them both. Further it reveals the process of the production of the notebooks. Shows the previous state of the materials.

@@

These are example placement. The exhibition event is not held. It is only decided. These notebooks are not only exhibited there, but also I will turn it to an workshop that continue to produce notebooks and show people to how to do it. Maybe they will join me in the production of notebooks. They can bring their items and transform all together. We can share our ideas and thoughts with each other.

%
%
\subsection{Website}
Finding a new place to the discarded items is one of the main purposes of this project. Not only it finds a place in people’s life again, but also it finds another place in the digital space.

>>>>

In this chapter thesis project and its development progress will be explained in detail in the light of what has been discussed in previous chapters. The project will be examined in various dimensions. The relationship will be established between the previously mentioned artworks and project. How this project is related to the discussed concepts and how it can be positioned among the previous works will be explained.

Firstly, I briefly state and describe my work. Secondly, the development of the project and experiments are explained. Lastly, the final body of the work with its parts is presented.

%****************************************
\section{The Statement}
Refuse is part of people’s consumption practices. It is a very common concept from urban areas to rural ones, from modern societies to ancient ones. It is not wanted, thrown away, kept away. It is considered as useless, disgusting, abject, unattractive, and harmful to the people and nature. Contrary to the common approach as \citet[7]{thompson1979rubbish}, author of the book \textit{Rubbish Theory}, stated \quotes{one man’s trash is another man’s treasure}.

@@

Through this project, I put all my effort to re-imagine, re-consider and re-think trash for seeking new possibilities and alternatives. As it is discarded, missed potentials were discovered. Strictly speaking, this project aims to find new potentials on trash.

The purpose is not to build an object that is produced from discarded material in order to watch it from distance. The important thing is to interact with it. It is already discarded and general behavior is to avoid from it. This work must change it. It must call the viewer to interact with it. In other words, trash must be accepted by them. This project asks people to readmit the trash to the their life. For this reason notebooks are given away free at various places where people visit frequently. By doing that once thrown materials would find a place with new purposes in the community.

%****************************************
\section{Development Progress of the Project}
In order to have an insight of this project and realize artistic intentions, it is important to look at the process that reveals the path to the final work and critical decisions that were necessary to the completion of the project.

@@

@@

% [SODA CAPS]
I remember from my childhood that we were collecting soda caps and were playing with them. We were looking everywhere to collect them. Some caps were found less and they worth more in the game. We put them on the railways to make them flat. After the train had passed we had perfect flat caps for our games. At that time, it was not waste for us. It has a value and part of our games and indispensable source of fun. The notion of playing with it can be related with the Picasso’s collage work \textit{Still Life with Chair Caning} mentioned previously. He pasted letters of the French word \quotes{JOU} means \quotes{game} signifying that playing with the materials. We were already playing with trash while we were child. Therefore, it can be better understand that why Picasso said that \quotes{every child is an artist}.

% [VASVIYE HOCA]
Another important memory is from my primary school teacher. When I was at third class, our teacher wanted us to bring colorful papers to make something (whatever it is I do not remember now). The day after we brought some colorful paper that was bought from stationery store. Our papers’ shape and color looked same. However her papers were different in every dimension. They were cut out from packages and advertisements. I remember clearly that she suggested us to do the same: \quotes{Do not throw out packages, look for the useful parts and keep them to use later.} Whenever I am going to throw something away, this comes to my mind: Can it be reused? Every time it is not obvious that how can it be useful, but at some point it becomes an \quotes{indispensable practice.} Further, this practice also supports to the \textit{bricolage} methodology. Keeping many things makes easier to build (or produce) whatever at hand.

@@

I am very uncomfortable with the act of discarding. I often think that by discarding objects, I am missing their potentials. I must find another use which is more effective than throwing away. At least I can give them to someone else who can use it for own purposes. Seeing papers on the waste bin is a disappointment for me. As artist Aaron Kramer argues and I strongly agree with that \quotes{[t]rash is the failure of imagination} \citep[as cited in][16]{meyer2007turning}. There is a lot of effort to produce objects and it seems that all this efforts are wasted. What I mention is not related about ultimate productivity. It is more close to being thoughtful, and taking responsibility of tools, items and objects that we are using. Rather than throwing out, creating a way that all are have chance to live together is much more close to my perception. On the contrary, sometimes I became overwhelmed because of the significant amount of leftover from daily activities and in order to move on I need to leave them behind. It is the result of inevitable rhythm of modern life style. Consuming and wasting is part of people’s daily routine and most of the time they are not aware of what they consume, and what they discard. In the scope of this thesis subject, saving objects rather them getting rid of them is an effective way of breaking the routines. The usage of that object can be decided in the future.

% [encouraging point]
One of the encouraging factor of working on trash is to explore something different that others cannot see. In other words, it contains eureka moments. Because it is ignored by discarding and therefore it has undiscovered potentials. It waits us to be discovered and possibly (as Picasso’s explanation on his work \textit{Bull’s Head}\footnote{See epigraph in the beginning of chapter 3}, 1942) discovered again and again.

% [EXPERIMENTS AT HOME]
My remarkable interaction with trash is from recent years. Long years I stayed at dormitory during when I was high school and university while undergraduate. Dormitory is place that living spaces is shared with many people. Therefore there is a limited place for you and your possessions and there is no space for unused items. After I graduated from university I started living in a rented house. This gives me to opportunity to save the materials to use them later. My very first intention is to use them on my paintings. Instead of white blank canvas I have experimented with different surfaces and I have collected boxes and cart boards from my friends. My aim is to turn discard materials to the different things. Here are some of my previous works.

@@

After turning discarded materials to different uses I began to collect them more eagerly. At the same time I also gave what I did to my friends as a birthday gift. I was experimenting with the material and at some point, with the help of my mother, I made my first notebook covered by a part of Starbucks bag. Papers inside of it are blank mimeograph paper which different than white paper. Particularly purpose is to move away from white paper because I feel uncomfortable when using it. It looks to me very bright. Therefore mimeograph paper is better for me. At that time it was a great victory for me. I started to carry it and use it every possible place. I show them to people and some of them supported to me giving their discard. One of them is my supervisor. He started to collect packages for me. After turning his packages to notebooks, he suggested me to go deeper in this subject and make my thesis project from them. Thereafter what I need is more discarded material.

@@

%****************************************
\subsection{Collecting}
Collecting discarded objects has carried out through the efforts of mine and people around me. Lars Eighner’s instructions helped me a lot when collecting objects. He writes that:

\begin{quote}
Eating safely from the Dumpsters involves three principles, using the senses and common sense to evaluate the condition of the found materials, knowing the Dumpsters of a given area and checking them regularly, and seeking always to answer the question \quotes{Why was this discarded?} \citep[as cited in][6]{strasser1999waste}.
\end{quote}

@@

Moreover there are other rules that define the frame of collecting act. Paper or package must be trash or will be trashed. For this project anything must not be intentionally bought to use as trash. In other words first rule is to consume less and generate less trash. Lots of trash is already produced every time everywhere. Aim is to reach them.

% OWNER
\quotes{The question of who owns these discards is not trivial} \citep{zimring2012encyclopedia}. This question makes me feel uncomfortable because of thinking that stealing that someone elses’ objects. Particularly I felt it when collecting left out papers from Bilkent Computer Laboratory. You already know that it owned by somebody else and after a while owner can take them. I solved this issue checking the leaved papers in every couples hours. If it stays there I take them. Further I encountered that If nobody takes them for a long time servants collect them.

@@

% CONTRIBUTIONS OF PEOPLE
As mentioned previously many people contributed to the project by collecting. Whenever I meet my friends I ask them to save their discard for the project. The importance of collaboration with them can be understood the term \quotes{citizen science} which is \quotes{defined as scientific research performed in part or in whole by volunteers who are not professional scientists} \citep{robson2012using}. I need to note that this is not a scientific project but the notion belongs there. The value of it lies in the collective intelligence of the contributors. In the scope of this thesis the term can be interpreted as people bring their own creativity and interpretation to the project by collecting the items that they reach and select. When the number of contributors are increased the diversity of the collected material will be increased and this will echo in the notebooks.

It can be viewed as a co-authored work. Because many people helped me to collect the papers and also many people leave their trail to the paper and also they bring meaning to the them. After production of papers I also want them to continue as co-authored process. In other words many people will be part of this project. As people see different things on objects and bring them to me. the variety of things is increased in time. Because people said me that \quotes{I thought that you can use it. Do you collect also these?}. They bring their unique approach.

@@

% What type of activities generates waste.
Beyond social status it reveals activities that generate such a waste.It can be also extracted that most of the trashed papers are thrown away from feeding activity. On the other hand vastness of some types of disposables reveal the dark side of enterprises such as Starbucks and Burger King. Few people thought that after seeing the project I only collected Starbucks packets. There is no such a limitation but as a result of my habitat the number of them suppress the others.

@@

Collected items (Figure \ref{fig:CollectedAllTogether}) are stored in my room for a long time. Most of them do not belong to me. Here the detailed information of the collected materials including the place where it is collected and their previous usage are listed.

\textbf{Burger King packages} (Figure \ref{fig:BurgerKingBag}) are collected from the office where I work. At lunch time, some of the colleagues order their meal from Burger King which is a global chain of fast food restaurant. Meals are delivered to the office with paper packages. The package of the meal is collected and used as a cover of notebooks.

Inside the bag one paper package is used for French fries which is oily and paper is very good to absorb oil. Some people think that it is disgusting to use it for other purposes because of oil onto it. But immediately I replied them \quotes{you just eat that oily fires so what is wrong with it?} At the same time, what \citet[5]{strasser1999waste} said comes to my mind; \quotes{Nothing is inherently trash}. She refers to resolution on dirt of \citet[36]{douglas2003purity}:

@@

\textbf{Starbucks packages} (Figure \ref{fig:StarbucksBag}) are collected from my supervisor, office where I work and my friend. Particularly my friend saved and collected them from her friends. Few of them collected from waste bins. They are used as a cover of notebooks. There are different types of packets from this category. As claimed by the company all of them are built from recycled ---precisely downcycled--- paper. There is also one category of object that is not paper. It is package of coffee beans. It is airtight and waterproof. It is very solid and strong. It has well graphic designs with different color and illustration.

\textbf{Modshifters papers} (Figure \ref{fig:Modshifters}) are used to cover the dinner table. Beginning of the table there is a big roll of paper to cover the table. At the end of the day used parts are cut out and thrown away. One of my visit I stayed there at a late time and catch the waiter when collecting the used papers and preparing tables to the next customer. I asked kindly him to give me these papers and he accepted. They are used as a cover of notebooks. All of them contains track of people who sit down there. I do not know any of them and I am not sure that the paper that they are laid down on their dinner turned to such a thing.

@@

\textbf{Order slips} (Figure \ref{fig:BankSlips}) are collected from bank office located at Technopolis of Middle East Technical University. They are used to record the order of customers. I realized that back of them are empty and appropriate for writing. Further there many others are waiting on the bin. After this experiment I really convinced that I am really looking for trash everywhere every time. The act of collecting spread through the every moment of life.

@@

%****************************************
\subsection{Transformation}
At one side while paper is being collected, on the other side I started to make experiments with the material. I questioned what else can be made with these collected materials and tried to cover (wrap out) objects and plants in the public. However these do not give the result what I want. People cannot interact with these and possibly they will turn to trash again. For that reason the first approach which is making notebooks are more appropriate to accomplish the purpose.

@@

% [notebooks]
There are two main parts of notebooks; cover and pages. Papers that are mostly blank (empty) are used inside of the notebooks for writing. Others that are solid and colorful are used as a cover for notebooks. Mixing (juxtaposing) materials is main method in the production process. I combine different papers and packages. I try to generate every possible combination. The purpose is to create as many alternatives as I can because trash is ignored and people do not give it alternatives. Therefore these notebooks provide many alternatives to show that otherwise is possible.

@@

Size of the paper used inside of the notebooks are mostly equal to the size of an A4 sheet (dimensions: 210mm x 297mm). There are different dimensions of notebooks. Most of them are half of A4 sheet or quarter of it. Collected papers' dimension shape the size of the notebooks. Usage of papers differs among the notebooks. Especially for the papers taken from Varuna Gezgin written part is folded inward by hiding it. However for the ones collected from Bilkent Computer Laboratory written side of it is left open. Their content comes from various students whose departments are also various. In other words, they offer great diversity and they are mixed to allow different interpretations. Further they provide what students from other departments work on.

@@

While transforming collected materials my method can be explained with the term upcycling mentioned at the end of the chapter 1. Further I tried to use objects already at hand. In other words I followed the \textit{bricolage} technique. However I have to do buy some tools for my work. Therefore it is hard to consider it as \textit{bricolage} but my intention is to follow it.

Moreover I have reviewed other notebook projects. Notable ones are \textit{The Sketchbook Project} and \textit{myDetour} from Moleskine. In these projects anyone who bought the notebooks from them can participate and send their notebooks filled with illustrations and paintings. These projects can be seen as crowd-sourced library. Further the pages of the notebooks are available at online. After checking sketches in these notebooks, the diversity of them surprised me a lot. It forced me think that the transformation process is not limited with the production of notebooks from trash. After notebooks are given to the people, their transformation will continue. Therefore I think the process of transformation is a continuous act.

@@

%****************************************
\subsection{Demonstration}
After producing notebooks the most important point is to present and reach to people. Main question is how to give them to people. Starting point is to stack the notebooks onto the each other, and there are other ideas like putting them inside a waste bin to re-do the action of throwing away. Afterwards somehow they should understand that these notebooks can be taken. I set up models and tried some installations of them.

@@

They were early interpretations and ideas. I imagined an exhibition environment to place notebooks (Figure \ref{fig:exhibition}). Different placements were tried. All of them looked like a sculpture. I have inserted no special meaning as sculpture to this objects. I also afraid that people will not interact with them and look them away. My purpose is give people. Aim is to spread the idea by making something useful from trash and to increase diversity, to encourage people in order to embrace the trash.

@@

To hold notebooks together I made several small boxes from discarded card boxes. Notebooks were put inside of them. In the production phase of the boxes, I discovered that my folded notebooks are fit into the many card boxes. In other words dimensions of them are not very different. This property of items make my production more easier. It shows that industrial system supports the reuse of the object with less effort in this way.

@@

%****************************************
\section{Parts of the Final Work}
As also can be understood by the previous section the project introduced in this thesis have different parts. Each part support the other. They are connected with each other. Different parts lead different inquiries.

%****************************************
\subsection{Notebooks}
Notebooks have been used for different purposes such as drawing, writing and recording, ideas or memorandum by many artists, scientists, and thinkers through the ages. Further the same exists for me and people around me. Although there are digital alternatives of it paper maintains its place in the community.

@@

% Meaning of stories behind the process
There are different stories behind the objects. I try to record their stories (by photography and taking notes) but some of them were not possible. However I still use them in the work because even if I missed their story, with their materiality their history reveal themselves. They still have histories but they need to be rewritten. Maybe forgetting all the history maybe creates different notebooks.

@@

Giving name is another hard topic for me. However point out the story behind the notebooks and properties of it. These names gives clues about how they are combined what they represent, defines the aspects of them. Name of the paper can be related with this place and related with the meaning. These names captures the memory of notebooks. However for one series I borrowed the idea of nonsense word generation from the Kurt Schwitters. Similar to the the production of the nonsense word \textit{Merz} I decided to name some of the notebooks a nonsense word. There are notebook series that have a name. Here are the name of notebooks and their story:

@@

\textbf{Daro.} For some of the notebooks it is not obvious to name them. At this point approach of Kurt Schwitters provides me a way. He generated his frequently used word \textit{Merz} from an advertisement. It is a nonsense word and used in most of his works. I followed what he did and generated a word from the papers that used inside of the notebooks. I use the word \textit{Daro} extracted from \textit{Vardaroma Dondurmacısı} for the ones whose story are not obvious and significant. The word found one the of papers collected from near the printer inside of computer laboratory.

%****************************************
\subsection{Exhibition}
Especially exhibition of notebooks is the most hard topic in the development of project. To find an appropriate place for the notebooks is a great challenge for me. Main idea is here to find a place that can also be considered as an alternative art space because in every dimension alternative is seek through the project. It must not only fit inside of white cube and must reach to audience. Therefore frequently visited places is a perfect locations for the project.

Bilkent library and computer laboratory are one of the places where notebooks are placed inside of boxes made from card boards (Figure \ref{fig:BilkentLibrary}-\ref{fig:BilkentLaboratory}).  Libraries are places where books are reused and shared among many people. It can be viewed as a place of sharing culture. Further it is a place that students from several department visit the same location. Particularly in Bilkent there are small papers to write down the shelve numbers of books. It is a great place to leave notebooks there because people already need something to write. They can find notebooks when they need them. For the computer laboratory case, the main purpose is to provide an alternative near the place where mostly papers are carelessly wasted.

\begin{figure}[!tbp]
  \begin{minipage}[b]{0.48\textwidth}
    \includegraphics[width=\textwidth]{project_graphics/bilkent1.jpg}
    \caption{Sample placement at Bilkent Library}
    \label{fig:BilkentLibrary}
  \end{minipage}
  \hfill
  \begin{minipage}[b]{0.48\textwidth}
    \includegraphics[width=\textwidth]{project_graphics/bilkent2.jpg}
    \caption{Sample placement at Bilkent Computer Laboratory}
    \label{fig:BilkentLaboratory}
  \end{minipage}
\end{figure}

@@

% Torun
Through this project another exhibition alternative is thought. In the notable museums and art galleries there are gift shops that people can buy souvenirs. Often they are mass produced imitations and replicas of well-known artists work embedded various objects such as cover on the notebooks. At this point I image an art space you can get these types of items freely. In other words alternative to the existing approach this types of gadgets. Another alternative interpretation can be giving away different that the industrial items such as notebooks from trash as in my project. For this type of approach Torun is a great place to show my notebooks and give them away. Torun describes itself as a place for free place for sharing art. The place is open to everybody and makes open call to any artist. There are no security cameras on this gallery therefore no one is watching you at there. The project is shared with Torun and it is accepted.

I used wooden market boxes to lift up them from ground to more reachable height. It is hole body of stand to put notebooks. Notebooks are placed on the top box and below them I just put the collected items to provide for those who want to make their own notebook. People can get them both. Further it reveals the process of the production of the notebooks and shows the previous state of the materials.

@@

Figure \ref{fig:SamplePlacementAtTorun} shows the example placements in Torun. The exhibition event is not held. It is only decided. These notebooks will not only be exhibited there, but also I will turn it to an workshop that I continue to produce notebooks and show people to how to do it. Maybe they will join me in the production of notebooks. They can bring their items and we can transform them all together. We can share our ideas and thoughts with each other.

%****************************************
\subsection{Website}
Finding a new place to the discarded items is one of the main purposes of this project. Not only it finds a place in people’s life again, but also it finds another place in the digital space.
